---
title: "DEP analysis"
author: "Rui Nascimento"
date: "19/01/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# DEP Analysis Test

```{r, message=FALSE}
library("DEP")
library("dplyr")
library("stringr")
library("rentrez")
library("org.Vvinifera.eg.db")
library("AnnotationDbi")
library("topGO")
library("openxlsx")
library("data.table")
```


## Label-Free Quantification

```{r}
data <- read.delim("Vitis_Protein_Groups.tsv", header=TRUE, sep="\t", stringsAsFactors = FALSE)
dim(data)
data <- filter(data, Reverse != "+", Potential.contaminant != "+")
dim(data)
```

Are any there duplicated genes?

```{r}
data$Protein.IDs %>% duplicated() %>% any()
```

Define functions for later use

```{r}
getGI <- function(s) {
  ## Helper function to extract GI from fasta headers in MAxQuant results
  ex <- str_extract_all(string = s, pattern = "gi\\|[:digit:]+\\|")[[1]]
  str_extract_all(string = ex, pattern = "[:digit:]+",simplify = T)[,1]
}

get1GI <- function(x){
 # Get 1 GI, priority to GI with XP annotation, otherwise get the first GI 
  if (str_detect(x, "XP")) {
    gi <- str_extract(string = x, pattern = "(?<=gi\\|)[:number:]+(?=\\|ref\\|XP)")
  } else {
      gi <- str_extract(string = x, pattern = "(?<=gi\\|)[:number:]+")
    }
  return(gi)
}

protein2gene <- function(x, api_key = NULL){
  if (is.null(api_key)) {
  gene_gi <- entrez_link(dbfrom="protein", db="gene", id = x)$links$protein_gene[1]
  } else {
  gene_gi <- entrez_link(dbfrom="protein", db="gene", id = x, api_key = api_key)$links$protein_gene[1]
  }
  if (is.null(gene_gi)){gene_gi <- ""}
  return(gene_gi[[1]])
}

# Get protein ID Name
proteinName <- function(x, api_key = NULL){
  if (is.null(api_key)) {
  protName <- entrez_summary(db = "protein", id = x)$title
  } else {
  protName <- entrez_summary(db = "protein", id = x, api_key = api_key)$title
  }
  if (is.null(protName)){protName <- ""}
  return(protName)
}

# Get all names
get_names <- function(prot_ids) {
  gi <- getGI(prot_ids)
  name <-  lapply(gi, function(id) (entrez_summary(id = id, db = "protein")$title))
  return(paste(name, collapse = ";"))
}

```


```{r}
#Get 1 GI and add to ID column
data$ID <- as.character(sapply(data$Protein.IDs, get1GI))
data$name <- as.character(data$Protein.IDs)
```


```{r}
data_unique <- make_unique(data, names = "name", ids = "ID", delim = ";")
```


## Generate a SummarizedExperiment object using an experimental design

Choose 6 and 12 hpi timepoints

```{r}
LFQ_columns <- grep("Intensity\\.R(m|i)(6|12)", colnames(data)) # get LFQ column numbers
experimental_design <- read.delim("ExperimentalDesign_no24.csv", header=TRUE, sep="\t", stringsAsFactors = FALSE)
data_se <- make_se(data, LFQ_columns, experimental_design)
```


```{r}
plot_frequency(data_se)
```


## Filter missing values

Filter for proteins that are identified in 2 out of 3 replicates of at least one condition

```{r}
data_filt <- filter_missval(data_se, thr = 1)
```

Barplot of the number of identified proteins per samples

```{r}
plot_numbers(data_filt)
```

Barplot of the protein identification overlap between sample

```{r}
plot_coverage(data_filt)
```

## Normalization

Normalize the data
```{r}
data_norm <- normalize_vsn(data_filt)
```

Visualize Normalization

```{r}
plot_normalization(data_filt, data_norm)
```

## Missing values imputation

View missing values

```{r}
plot_missval(data_filt)
```

Plot intensity distributions and cumulative fraction of proteins with and without missing values
```{r}
plot_detect(data_filt)
```

Impute missing data using random draws from a Gaussian distribution centered around a minimal value (for MNAR)

```{r, message=FALSE}
data_imp <- impute(data_norm, fun = "MinProb", q = 0.01)
```

Intensity distributions before and after imputation

```{r}
plot_imputation(data_norm, data_imp)
```

## DE analysis Global

```{r}
data_diff_manual <- test_diff(data_imp, type = "manual", 
                              test = c("Ri6h_vs_Rm6h", "Ri12h_vs_Rm12h"))
```

```{r}
dep <- add_rejections(data_diff_manual, alpha = 1, lfc = log2(1.5))
```
  
## Results Table

```{r}
# Generate a results table
data_results <- get_results(dep)

# Number of significant proteins
data_results %>% filter(significant) %>% nrow()
```

Get Gene IDs from Protein IDs from NCBI servers

```{r}
api_key <- readLines("~/.ncbi_api")
data_results$GENE_GI <- sapply(data_results$ID, protein2gene, api_key = api_key, simplify = TRUE, USE.NAMES = FALSE)
data_results$PROTEIN_NAME <- sapply(data_results$ID, proteinName, api_key = api_key, simplify = TRUE, USE.NAMES = FALSE)
```


```{r}
plot_pca(dep, x = 1, y = 2, n = 500, point_size = 4)
```

## 6 hpi

### PCA

```{r}
plot_pca(dep[, c(4,5,6,10,11,12)], x = 1, y = 2, n = 500, point_size = 4)

```

### Go enrichment analysis

Genererate genelist of differentially accomulated proteins using the Gene IDs

```{r}
geneList6h <- data_results$Ri6h_vs_Rm6h_p.val
names(geneList6h) <- data_results$GENE_GI
head(geneList6h)
```

Run the Go enrichment tests

```{r}
GOdata6h <- new("topGOdata",
              ontology = "BP",
              allGenes = geneList6h,
              geneSelectionFun = function(x) (x < 0.01),
              annot = annFUN.org , mapping = "org.Vvinifera.eg.db",
              ID = "entrez")

resultKS6h <- runTest(GOdata6h, algorithm = "weight01", statistic = "fisher")

tab6h <- GenTable(GOdata6h, raw.p.value = resultKS6h, topNodes = length(resultKS6h@score), numChar = 120)
# Convert p.value to numeric
tab6h$raw.p.value <- as.numeric(tab6h$raw.p.value)

head(tab6h, 15)

par(cex = 0.6)
showSigOfNodes(GOdata6h, score(resultKS6h), firstSigNodes = 5, useInfo = "def")
par(cex = 1)
```


## 12 hpi

```{r}
plot_pca(dep[, c(1,2,3,7,8,9)], x = 1, y = 2, n = 500, point_size = 4)
```

Genererate genelist of differentially accomulated proteins using the Gene IDs

```{r}
geneList12h <- data_results$Ri12h_vs_Rm12h_p.val
names(geneList12h) <- data_results$GENE_GI
head(geneList12h)
```

Run the Go enrichment tests

```{r}
GOdata12h <- new("topGOdata",
                ontology = "BP",
                allGenes = geneList12h,
                geneSelectionFun = function(x) (x < 0.01),
                annot = annFUN.org , mapping = "org.Vvinifera.eg.db",
                ID = "entrez")

resultKS12h <- runTest(GOdata12h, algorithm = "weight01", statistic = "fisher")

tab12h <- GenTable(GOdata12h, raw.p.value = resultKS12h, topNodes = length(resultKS12h@score), numChar = 120)
# Convert p.value to numeric
tab12h$raw.p.value <- as.numeric(tab12h$raw.p.value)

head(tab12h, 15)

par(cex = 0.6)
showSigOfNodes(GOdata12h, score(resultKS12h), firstSigNodes = 5, useInfo = "def")
par(cex = 1)
```

## Save Results

Save GO results with protein data.

Define FoldChange to highlight in excel.

```{r}
FC <- 1.5
upreg <- paste0(">", log2(FC))
downreg <- paste0("<-", log2(FC))
```

Set Styles.

```{r}
negStyle <- createStyle(fontColour = "#9C0006", bgFill = "#FFC7CE")
posStyle <- createStyle(fontColour = "#006100", bgFill = "#C6EFCE")
```

Save 6hpi GO Results

Create excel workbook to put data in.

```{r}
fileName <- paste0("results/6hpi_", "go_results.xlsx")
excel <- createWorkbook()
```

Create the Summary Page.

```{r}
addWorksheet(excel, "Summary")
writeData(tab6h[tab6h$raw.p.value < 0.05, ], wb = excel, sheet = "Summary", rowNames = FALSE)
```

For loop to generate the diferent dataframes and save to a list.

```{r, warning=FALSE, message=FALSE}
GoDF <- list()
for (x in 1:sum(tab6h$raw.p.value < 0.05)) {
  goID <- tab6h[x, "GO.ID"]
  ids <- genesInTerm(GOdata6h, goID)
  annot <- data.table(AnnotationDbi::select(org.Vvinifera.eg.db, keys=ids[[1]], column=c("symbol", "gene_name"), keytype="gene_id"))
  tmp <- data.table(data_results)
  expVals <- tmp[GENE_GI %in% annot$gene_id, ]
  final_annot <- merge.data.table(annot, expVals, by.x ="gene_id", by.y = "GENE_GI")
  setorder(final_annot, Ri6h_vs_Rm6h_p.adj)
  GoDF[[goID]] <- final_annot
}
```

For loop to add every sheet, every sheet is a GO_ID with a p.value < 0.05.
Add Hyperlinks to Summary Page.
Add conditional formating to highligh significant foldchange / p.values.

```{r}
for (x in 1:sum(tab6h$raw.p.value < 0.05)) {
  goID <- tab6h[x, "GO.ID"]
  sheetname <- substring(tab6h[x, "GO.ID"], 4)
  addWorksheet(excel, sheetname)
  writeData(GoDF[[x]], wb = excel, sheet = sheetname, rowNames = FALSE)
  writeFormula(excel, "Summary", startRow = x+1, 
               x = makeHyperlinkString(sheet = sheetname, row = 1, col = 1
                                       , text = goID))
  conditionalFormatting(excel, sheetname, cols=14, rows=2:(nrow(GoDF[[x]])+1), rule=upreg, style = posStyle)
  conditionalFormatting(excel, sheetname, cols=14, rows=2:(nrow(GoDF[[x]])+1), rule=downreg, style = posStyle)
  conditionalFormatting(excel, sheetname, cols=7, rows=2:(nrow(GoDF[[x]])+1), rule="<0.01", style = posStyle)
  conditionalFormatting(excel, sheetname, cols=9, rows=2:(nrow(GoDF[[x]])+1), rule="<0.05", style = posStyle)
}

saveWorkbook(excel, file = fileName, overwrite = TRUE)
```

Save 12hpi GO Results

Create excel workbook to put data in.

```{r}
fileName <- paste0("results/12hpi_", "go_results.xlsx")
excel <- createWorkbook()
```

Create the Summary Page.

```{r}
addWorksheet(excel, "Summary")
writeData(tab12h[tab12h$raw.p.value < 0.05, ], wb = excel, sheet = "Summary", rowNames = FALSE)
```

For loop to generate the diferent dataframes and save to a list.

```{r, warning=FALSE, message=FALSE}
GoDF <- list()
for (x in 1:sum(tab12h$raw.p.value < 0.05)) {
  goID <- tab12h[x, "GO.ID"]
  ids <- genesInTerm(GOdata12h, goID)
  annot <- data.table(AnnotationDbi::select(org.Vvinifera.eg.db, keys=ids[[1]], column=c("symbol", "gene_name"), keytype="gene_id"))
  tmp <- data.table(data_results)
  expVals <- tmp[GENE_GI %in% annot$gene_id, ]
  final_annot <- merge.data.table(annot, expVals, by.x ="gene_id", by.y = "GENE_GI")
  setorder(final_annot, Ri12h_vs_Rm12h_p.adj)
  GoDF[[goID]] <- final_annot
}
```

For loop to add every sheet, every sheet is a GO_ID with a p.value < 0.05.
Add Hyperlinks to Summary Page.
Add conditional formating to highligh significant foldchange / p.values.

```{r}
for (x in 1:sum(tab12h$raw.p.value < 0.05)) {
  goID <- tab12h[x, "GO.ID"]
  sheetname <- substring(tab12h[x, "GO.ID"], 4)
  addWorksheet(excel, sheetname)
  writeData(GoDF[[x]], wb = excel, sheet = sheetname, rowNames = FALSE)
  writeFormula(excel, "Summary", startRow = x+1, 
               x = makeHyperlinkString(sheet = sheetname, row = 1, col = 1
                                       , text = goID))
  conditionalFormatting(excel, sheetname, cols=13, rows=2:(nrow(GoDF[[x]])+1), rule=upreg, style = posStyle)
  conditionalFormatting(excel, sheetname, cols=13, rows=2:(nrow(GoDF[[x]])+1), rule=downreg, style = posStyle)
  conditionalFormatting(excel, sheetname, cols=6, rows=2:(nrow(GoDF[[x]])+1), rule="<0.05", style = posStyle)
  conditionalFormatting(excel, sheetname, cols=8, rows=2:(nrow(GoDF[[x]])+1), rule="<0.05", style = posStyle)
}

saveWorkbook(excel, file = fileName, overwrite = TRUE)
```


```{r}
# Generate a wide data.frame
df_wide <- get_df_wide(dep)
# Generate a long data.frame
df_long <- get_df_long(dep)

write.table(df_wide, file = "results_wide.tsv", sep = "\t", row.names = FALSE)
write.table(df_long, file = "results_long.tsv", sep = "\t", row.names = FALSE)
write.table(data_results, "Results.tsv", sep = "\t", row.names = FALSE)
```

