---
title: "Grapevine – downy mildew rendez-vous: first hours’ proteomics of the incompatible interaction"
author: "Rui Nascimento"
date: "`r format(Sys.time(), '%d %B, %Y (%X)')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Enviroment Setup

Load Required Libraries

```{r, message=FALSE}
library("DEP")
library("dplyr")
library("plyr")
library("stringr")
library("rentrez")
library("org.Vvinifera.eg.db")
library("AnnotationDbi")
library("topGO")
library("openxlsx")
library("data.table")
library("ggplot2")
library("reshape2")
library("RColorBrewer")
library("ggthemes")
library("rrvgo")
library("DT")
library("ComplexHeatmap")
library("ggpubr")
```

Define some functions for later use

```{r}
# Helper function to extract GI from fasta headers in MaxQuant results
getGI <- function(s) {
  ex <- str_extract_all(string = s, pattern = "gi\\|[:digit:]+\\|")[[1]]
  str_extract_all(string = ex, pattern = "[:digit:]+",simplify = T)[,1]}

# Get 1 GI, priority to GI with XP annotation, otherwise get the first GI 
get1GI <- function(x){
  if (str_detect(x, "XP")) {gi <- str_extract(string = x, pattern = "(?<=gi\\|)[:number:]+(?=\\|ref\\|XP)")}
  else {gi <- str_extract(string = x, pattern = "(?<=gi\\|)[:number:]+")}
  return(gi)}

# Get the Gene ID of a protein
protein2gene <- function(x, api_key = NULL){
  if (is.null(api_key)) {gene_gi <- entrez_link(dbfrom="protein", db="gene", id = x)$links$protein_gene[1]}
  else {gene_gi <- entrez_link(dbfrom="protein", db="gene", id = x, api_key = api_key)$links$protein_gene[1]}
  if (is.null(gene_gi)){gene_gi <- ""}
  return(gene_gi[[1]])}

# Get protein ID Name
proteinName <- function(x, api_key = NULL){
  if (is.null(api_key)) {protName <- entrez_summary(db = "protein", id = x)$title}
  else {protName <- entrez_summary(db = "protein", id = x, api_key = api_key)$title}
  if (is.null(protName)){protName <- ""}
  return(protName)}

# Get Protein Code, priority to XP annotation, otherwise get other 
getProteinCode <- function(x){
  if (str_detect(x, "XP")) {gi <- str_extract(string = x, pattern = "XP_[:number:]+(\\.[:number:])*")}
  else {gi <- str_extract(string = x, pattern = "(?<=gi\\|[:alnum:]{6,9}\\|[:alnum:]{1,4}\\|)[:alnum:]+_*[:alnum:]+(\\.[:number:])*")}
  return(gi)}

# Get all names
get_names <- function(prot_ids) {
  gi <- getGI(prot_ids)
  name <-  lapply(gi, function(id) (entrez_summary(id = id, db = "protein")$title))
  return(paste(name, collapse = ";"))}

# Display table function for RMarkdown
DTable <- function(dt){
  datatable(dt, extensions = 'Buttons', options = list(
    dom = 'Blfrtip',
    pageLength = 5,
    lengthMenu = c(5, 10, 20, 50, 100),
    buttons = c('copy', 'csv', 'excel')))}

# Catitalize first letter of every word in a string
CapStr <- function(y) {
  c <- strsplit(y, " ")[[1]]
  paste(toupper(substring(c, 1,1)), substring(c, 2),
      sep="", collapse=" ")
}
```

# Label-Free Quantification

Read MaxQuant Protein Groups file and remove Reverse hits and contaminants

```{r}
data <- read.delim("Vitis_Protein_Groups.tsv", header=TRUE, sep="\t", stringsAsFactors = FALSE)
dim(data)
data <- filter(data, Reverse != "+", Potential.contaminant != "+")
dim(data)
```

Are any there duplicated genes?

```{r}
data$Protein.IDs %>% duplicated() %>% any()
```

Add GeneID info to dataframe

```{r}
#Get 1 GI and add to ID column
data$ID <- as.character(sapply(data$Protein.IDs, get1GI))
data$name <- as.character(data$Protein.IDs)
```

Generate unique identifiers based on "name" and "id" columns.

```{r}
data_unique <- make_unique(data, names = "name", ids = "ID", delim = ";")
```

Generate a SummarizedExperiment object using column names

```{r}
LFQ_columns <- grep("Intensity\\.R(m|i)(6|12)", colnames(data)) # get LFQ column numbers
experimental_design <- read.delim("ExperimentalDesign_no24.csv", header=TRUE, sep="\t", stringsAsFactors = FALSE)
data_se <- make_se(data, LFQ_columns, experimental_design)
```

Frequency Plot

```{r}
plot_frequency(data_se)
```

## Filter missing values

Filter for proteins that are identified in 2 out of 3 replicates of at least one condition

```{r}
data_filt <- filter_missval(data_se, thr = 1)
```

Barplot of the number of identified proteins per samples

```{r}
plot_numbers(data_filt)
```

## Normalization

Normalize the data
```{r}
data_norm <- normalize_vsn(data_filt)
```

Visualize Normalization

```{r}
plot_normalization(data_filt, data_norm)
```

## Missing values imputation

Impute missing data using random draws from a Gaussian distribution centered around a minimal value (for MNAR)

```{r, message=FALSE}
data_imp <- impute(data_norm, fun = "MinProb", q = 0.01)
```

## Differential accomulation analysis

Generate contrast tests for 6 and 12 hpi

```{r}
data_diff_manual <- test_diff(data_imp, type = "manual", 
                              test = c("Ri6h_vs_Rm6h", "Ri12h_vs_Rm12h"))
```

```{r}
dep <- add_rejections(data_diff_manual, alpha = 0.05, lfc = 1)
```
  
## Results Table

```{r}
# Generate a results table
data_results <- get_results(dep)
```

Get Gene IDs from Protein IDs from NCBI servers

```{r}
api_key <- readLines("~/.ncbi_api")
data_results$GENE_GI <- sapply(data_results$ID, protein2gene, api_key = api_key, simplify = TRUE, USE.NAMES = FALSE)
data_results$PROTEIN_NAME <- sapply(data_results$ID, proteinName, api_key = api_key, simplify = TRUE, USE.NAMES = FALSE)
data_results$PROTEIN_ID <- sapply(data_results$name, getProteinCode, simplify = TRUE, USE.NAMES = FALSE)
```


```{r, message=FALSE}
plot_pca(dep, x = 1, y = 2, n = 100, point_size = 4)
```

### Heatmap of Fold-Change

```{r}
topX <- 50

heat <- data.table(data_results)[Ri12h_vs_Rm12h_p.val < 0.05 | Ri6h_vs_Rm6h_p.val < 0.05,c("PROTEIN_ID" ,"Ri12h_vs_Rm12h_ratio","Ri6h_vs_Rm6h_ratio")]
heat <- heat[, `:=` (Difference = abs(Ri12h_vs_Rm12h_ratio - Ri6h_vs_Rm6h_ratio))]
setorder(heat, -Difference)
heat <- heat[1:topX,-"Difference"]
heat <- as.matrix(heat[,-"PROTEIN_ID"], rownames=heat$PROTEIN_ID)
colnames(heat) <- c("6hpi", "12hpi")


heatmap <- ComplexHeatmap::Heatmap(heat,
                        name = "log2 Fold Change",
                        row_names_side = "left",
                        column_names_side = "top",
                        column_title = paste("Heatmap of top", topX, "Fold-Changes"),
                        row_title = "Protein ID",
                        row_title_rot = 0,
                        column_names_rot = 0,
                        column_names_centered = TRUE,
                        cluster_columns = FALSE)

pdf("fc_heatmap.pdf", width = 6, height = 10)
heatmap
dev.off()
heatmap
```


## 6 hpi

### PCA

PCA plot of the 100 most significat protein

```{r, message=FALSE}
plot_pca(dep[, c(4,5,6,10,11,12)], x = 1, y = 2, n = 100, point_size = 4)
```

### Differential Accumulation Results

6 hpi differential accumulation results.
Ratio is the 'log2(Fold)' change of the inoculated samples over the mock inoculated

```{r}
results_6hpi <- data.table(data_results)[,c("PROTEIN_NAME", "ID", "Ri6h_vs_Rm6h_ratio", "Ri6h_vs_Rm6h_p.val")]
DTable(results_6hpi)
```

## 12 hpi

### PCA

PCA plot of the 100 most significat protein

```{r, message=FALSE}
plot_pca(dep[, c(1,2,3,7,8,9)], x = 1, y = 2, n = 100, point_size = 4)
```

### Diferential Accumulation Results

12 hpi differential accumulation results.
Ratio is the 'log2(Fold)' change of the inoculated samples over the mock inoculated

```{r}
results_12hpi <- data.table(data_results)[,c("PROTEIN_NAME", "ID", "Ri12h_vs_Rm12h_ratio", "Ri12h_vs_Rm12h_p.val")]
DTable(results_6hpi)
```


# Go enrichment analysis

## 6hpi

Genererate genelist of differentially accomulated proteins using the Gene IDs

```{r}
geneList6h <- data_results$Ri6h_vs_Rm6h_p.val
names(geneList6h) <- data_results$GENE_GI
```

#### Biological Process

Run the Go enrichment tests

```{r, message=FALSE}
GOdata6h_BP <- new("topGOdata",
              ontology = "BP",
              allGenes = geneList6h,
              geneSelectionFun = function(x) (x < 0.05),
              annot = annFUN.org , mapping = "org.Vvinifera.eg.db",
              ID = "entrez",
              nodeSize = 3)

resultKS6h_BP <- runTest(GOdata6h_BP, algorithm = "classic", statistic = "fisher")

tab6h_BP <- GenTable(GOdata6h_BP, raw.p.value = resultKS6h_BP, topNodes = length(resultKS6h_BP@score), numChar = 120)
# Convert p.value to numeric
tab6h_BP$raw.p.value <- as.numeric(tab6h_BP$raw.p.value)

DTable(tab6h_BP)

par(cex = 0.6)
showSigOfNodes(GOdata6h_BP, score(resultKS6h_BP), firstSigNodes = 10, useInfo = "def")
par(cex = 1)
```

#### Molecular Function

Run the Go enrichment tests

```{r, message=FALSE}
GOdata6h_MF <- new("topGOdata",
              ontology = "MF",
              allGenes = geneList6h,
              geneSelectionFun = function(x) (x < 0.05),
              annot = annFUN.org , mapping = "org.Vvinifera.eg.db",
              ID = "entrez")

resultKS6h_MF <- runTest(GOdata6h_MF, algorithm = "classic", statistic = "fisher")

tab6h_MF <- GenTable(GOdata6h_MF, raw.p.value = resultKS6h_MF, topNodes = length(resultKS6h_MF@score), numChar = 120)
# Convert p.value to numeric
tab6h_MF$raw.p.value <- as.numeric(tab6h_MF$raw.p.value)

DTable(tab6h_MF)

par(cex = 0.6)
showSigOfNodes(GOdata6h_MF, score(resultKS6h_MF), firstSigNodes = 10, useInfo = "def")
par(cex = 1)
```


#### Cellular Component

Run the Go enrichment tests

```{r, message=FALSE}
GOdata6h_CC <- new("topGOdata",
              ontology = "CC",
              allGenes = geneList6h,
              geneSelectionFun = function(x) (x < 0.05),
              annot = annFUN.org , mapping = "org.Vvinifera.eg.db",
              ID = "entrez")

resultKS6h_CC <- runTest(GOdata6h_CC, algorithm = "classic", statistic = "fisher")

tab6h_CC <- GenTable(GOdata6h_CC, raw.p.value = resultKS6h_CC, topNodes = length(resultKS6h_CC@score), numChar = 120)
# Convert p.value to numeric
tab6h_CC$raw.p.value <- as.numeric(tab6h_CC$raw.p.value)

DTable(tab6h_CC)

par(cex = 0.6)
showSigOfNodes(GOdata6h_CC, score(resultKS6h_CC), firstSigNodes = 10, useInfo = "def")
par(cex = 1)
```

## 12hpi

Genererate genelist of differentially accomulated proteins using the Gene IDs

```{r}
geneList12h <- data_results$Ri12h_vs_Rm12h_p.val
names(geneList12h) <- data_results$GENE_GI
```

#### Biological Process

Run the Go enrichment tests

```{r, message=FALSE}
GOdata12h_BP <- new("topGOdata",
              ontology = "BP",
              allGenes = geneList12h,
              geneSelectionFun = function(x) (x < 0.05),
              annot = annFUN.org , mapping = "org.Vvinifera.eg.db",
              ID = "entrez")

resultKS12h_BP <- runTest(GOdata12h_BP, algorithm = "classic", statistic = "fisher")

tab12h_BP <- GenTable(GOdata12h_BP, raw.p.value = resultKS12h_BP, topNodes = length(resultKS12h_BP@score), numChar = 120)
# Convert p.value to numeric
tab12h_BP$raw.p.value <- as.numeric(tab12h_BP$raw.p.value)

DTable(tab12h_BP)

par(cex = 0.6)
showSigOfNodes(GOdata12h_BP, score(resultKS12h_BP), firstSigNodes = 10, useInfo = "def")
par(cex = 1)
```


#### Molecular Function

Run the Go enrichment tests

```{r, message=FALSE}
GOdata12h_MF <- new("topGOdata",
              ontology = "MF",
              allGenes = geneList12h,
              geneSelectionFun = function(x) (x < 0.05),
              annot = annFUN.org , mapping = "org.Vvinifera.eg.db",
              ID = "entrez")

resultKS12h_MF <- runTest(GOdata12h_MF, algorithm = "classic", statistic = "fisher")

tab12h_MF <- GenTable(GOdata12h_MF, raw.p.value = resultKS12h_MF, topNodes = length(resultKS12h_MF@score), numChar = 120)
# Convert p.value to numeric
tab12h_MF$raw.p.value <- as.numeric(tab12h_MF$raw.p.value)

DTable(tab12h_MF)

par(cex = 0.6)
showSigOfNodes(GOdata12h_MF, score(resultKS12h_MF), firstSigNodes = 10, useInfo = "def")
par(cex = 1)
```


#### Cellular Component

Run the Go enrichment tests

```{r, message=FALSE}
GOdata12h_CC <- new("topGOdata",
              ontology = "CC",
              allGenes = geneList12h,
              geneSelectionFun = function(x) (x < 0.05),
              annot = annFUN.org , mapping = "org.Vvinifera.eg.db",
              ID = "entrez")

resultKS12h_CC <- runTest(GOdata12h_CC, algorithm = "classic", statistic = "fisher")

tab12h_CC <- GenTable(GOdata12h_CC, raw.p.value = resultKS12h_CC, topNodes = length(resultKS12h_CC@score), numChar = 120)
# Convert p.value to numeric
tab12h_CC$raw.p.value <- as.numeric(tab12h_CC$raw.p.value)

DTable(tab12h_CC)

par(cex = 0.6)
showSigOfNodes(GOdata12h_CC, score(resultKS12h_CC), firstSigNodes = 10, useInfo = "def")
par(cex = 1)
```

Add child ontology to results dataframe for manual curation

```{r}
# Get ancestors given a GO id
ancestors <- function(go, ont){
  if (is.null(ont)) {ont <- Ontology(go)}
  if (ont == "BP") {res <- GOBPANCESTOR[[go]]}
  if (ont == "MF") {res <- GOMFANCESTOR[[go]]}
  if (ont == "CC") {res <- GOCCANCESTOR[[go]]}

  res
}

# Get GO ID given a gene id, also remove ancestors
get_ontology <- function(gi, metadat, ont) {
  ids <- metadat[gene_id == gi & ontology == ont, go_idall]
  anc <- c()
  for (go_id in ids) {
    anc <- append(anc, ancestors(go_id, ont=ont))
  }
  unique(anc)
  go_ids <- setdiff(ids, anc)
  
  return(paste(go_ids, Term(go_ids), collapse = "; "))
}

# Extract metadata information
metadat <- data.table(AnnotationDbi::select(org.Vvinifera.eg.db, keys=data_results$GENE_GI, column=c("go_idall"), keytype="gene_id"))
metadat$ontology <- sapply(metadat$go_idall, Ontology)
metadat$term <- sapply(metadat$go_idall, Term)

# Generate Ontology for the 3 GO Term
data_results2 <- data.table(data_results)
data_results2$BiologicalProcess <- sapply(data_results2$GENE_GI ,get_ontology, metadat=metadat, ont="BP")
data_results2$MolecularFunction <- sapply(data_results2$GENE_GI ,get_ontology, metadat=metadat, ont="MF")
data_results2$CelularComponent <- sapply(data_results2$GENE_GI ,get_ontology, metadat=metadat, ont="CC")
final_results <- data.table(data_results2)
write.csv(final_results, "Results.csv", row.names = FALSE)
```

### Reduce GO Semantic Similarity

6 hpi
```{r}
simMatrix6h_BP <- calculateSimMatrix(tab6h_BP[tab6h_BP$raw.p.value<0.05,c("GO.ID")],
                                orgdb="org.At.tair.db", ont="BP", method="Rel")
scores6h_BP <- setNames(-log10(tab6h_BP[tab6h_BP$raw.p.value<0.05,c("raw.p.value")]), tab6h_BP[tab6h_BP$raw.p.value<0.05,c("GO.ID")])
reducedTerms6h_BP <- reduceSimMatrix(simMatrix6h_BP, scores6h_BP, threshold=0.4, orgdb="org.At.tair.db")
treemapPlot(reducedTerms6h_BP)
write.csv(reducedTerms6h_BP, "SemanticReduction_6hpi_BP.csv", row.names = FALSE)
```

```{r}
simMatrix6h_MF <- calculateSimMatrix(tab6h_MF[tab6h_MF$raw.p.value<0.05,c("GO.ID")],
                                orgdb="org.At.tair.db", ont="MF", method="Rel")
scores6h_MF <- setNames(-log10(tab6h_MF[tab6h_MF$raw.p.value<0.05,c("raw.p.value")]), tab6h_MF[tab6h_MF$raw.p.value<0.05,c("GO.ID")])
reducedTerms6h_MF <- reduceSimMatrix(simMatrix6h_MF, scores6h_MF, threshold=0.4, orgdb="org.At.tair.db")
treemapPlot(reducedTerms6h_MF)
write.csv(reducedTerms6h_MF, "SemanticReduction_6hpi_MF.csv", row.names = FALSE)
```

```{r}
simMatrix6h_CC <- calculateSimMatrix(tab6h_CC[tab6h_CC$raw.p.value<0.05,c("GO.ID")],
                                orgdb="org.At.tair.db", ont="CC", method="Rel")
scores6h_CC <- setNames(-log10(tab6h_CC[tab6h_CC$raw.p.value<0.05,c("raw.p.value")]), tab6h_CC[tab6h_CC$raw.p.value<0.05,c("GO.ID")])
reducedTerms6h_CC <- reduceSimMatrix(simMatrix6h_CC, scores6h_CC, threshold=0.4, orgdb="org.At.tair.db")
treemapPlot(reducedTerms6h_CC)
write.csv(reducedTerms6h_CC, "SemanticReduction_6hpi_CC.csv", row.names = FALSE)
```

12 hpi

```{r}
simMatrix12h_BP <- calculateSimMatrix(tab12h_BP[tab12h_BP$raw.p.value<0.05,c("GO.ID")],
                                orgdb="org.At.tair.db", ont="BP", method="Rel")
scores12h_BP <- setNames(-log10(tab12h_BP[tab12h_BP$raw.p.value<0.05,c("raw.p.value")]), tab12h_BP[tab12h_BP$raw.p.value<0.05,c("GO.ID")])
reducedTerms12h_BP <- reduceSimMatrix(simMatrix12h_BP, scores12h_BP, threshold=0.4, orgdb="org.At.tair.db")
treemapPlot(reducedTerms12h_BP)
write.csv(reducedTerms12h_BP, "SemanticReduction_12hpi_BP.csv", row.names = FALSE)
```

```{r}
simMatrix12h_MF <- calculateSimMatrix(tab12h_MF[tab12h_MF$raw.p.value<0.05,c("GO.ID")],
                                orgdb="org.At.tair.db", ont="MF", method="Rel")
scores12h_MF <- setNames(-log10(tab12h_MF[tab12h_MF$raw.p.value<0.05,c("raw.p.value")]), tab12h_MF[tab12h_MF$raw.p.value<0.05,c("GO.ID")])
reducedTerms12h_MF <- reduceSimMatrix(simMatrix12h_MF, scores12h_MF, threshold=0.4, orgdb="org.At.tair.db")
treemapPlot(reducedTerms12h_MF)
write.csv(reducedTerms12h_MF, "SemanticReduction_12hpi_MF.csv", row.names = FALSE)
```

```{r}
simMatrix12h_CC <- calculateSimMatrix(tab12h_CC[tab12h_CC$raw.p.value<0.05,c("GO.ID")],
                                orgdb="org.At.tair.db", ont="CC", method="Rel")
scores12h_CC <- setNames(-log10(tab12h_CC[tab12h_CC$raw.p.value<0.05,c("raw.p.value")]), tab12h_CC[tab12h_CC$raw.p.value<0.05,c("GO.ID")])
reducedTerms12h_CC <- reduceSimMatrix(simMatrix12h_CC, scores12h_CC, threshold=0.4, orgdb="org.At.tair.db")
treemapPlot(reducedTerms12h_CC)
write.csv(reducedTerms12h_CC, "SemanticReduction_12hpi_CC.csv", row.names = FALSE)
```


## Plot GO results

Plot results with fold change associated with each signficant GO Term

```{r}
diffGO6h_BP <- data.table(reducedTerms6h_BP)[go == parent, go]

plot6h_full_BP <- data.table(tab6h_BP)[GO.ID %in% diffGO6h_BP]

plot6h_full_BP[,`:=`(Positive = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata6h_BP, x)[[1]] & Ri6h_vs_Rm6h_p.val<0.05 & Ri6h_vs_Rm6h_ratio > 2, Ri6h_vs_Rm6h_ratio])}, USE.NAMES = F), Negative = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata6h_BP, x)[[1]] & Ri6h_vs_Rm6h_p.val<0.05 & Ri6h_vs_Rm6h_ratio < -2, Ri6h_vs_Rm6h_ratio])}, USE.NAMES = F))]

plot6h_full_BP$Term2 <- sapply(plot6h_full_BP$Term, CapStr, USE.NAMES = F)


pal<-brewer.pal(3,"PuOr")
data <- plot6h_full_BP[raw.p.value<0.05 & Positive+Negative>3]

limit <- round_any(max(data[,Positive])+max(data[,Negative]),5,ceiling)
sep <- if (limit>30) {10} else {5}

data <- data.table::melt(data = data, id.vars = c("Term2"), measure.vars = c("Positive", "Negative"), variable.name = "Fold", value.name = "Count")
data[Fold == "Negative", `:=` (Col = pal[3])]
data[Fold == "Positive", `:=` (Col = pal[1])]

plotBP6h <- ggplot() + geom_bar(data=data, aes(x = Term2, y=Count, fill=Col), position="stack", stat="identity") +
  geom_hline(yintercept = 0, color =c("white")) +
  scale_fill_identity("Number of modulated proteins", labels = c("Up accumulated", "Down accumulated"), breaks=c(pal[1],pal[3]), guide="legend") + 
  theme_hc() + 
  coord_flip() +
  labs(title="6hpi GO Biological Process", y="",x="") +
  theme(plot.title = element_text(size=16, hjust=0.5)) +
  theme(axis.text = element_text(size=11)) +
  theme(axis.text.y = element_text(hjust=1)) +
  theme(legend.position = "bottom") +
  scale_y_continuous(breaks=seq(0,limit,sep), limits=c(0,limit)) +
  scale_x_discrete(label=function(x) stringr::str_trunc(x, 40))
plotBP6h
ggsave("BP6h.pdf", plot = plotBP6h, width = 10, height = 7, dpi = 600)
```


```{r}
diffGO6h_MF <- data.table(reducedTerms6h_MF)[go == parent, go]

plot6h_full_MF <- data.table(tab6h_MF)[GO.ID %in% diffGO6h_MF]

plot6h_full_MF[,`:=`(Positive = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata6h_MF, x)[[1]] & Ri6h_vs_Rm6h_p.val<0.05 & Ri6h_vs_Rm6h_ratio > 2, Ri6h_vs_Rm6h_ratio])}, USE.NAMES = F), Negative = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata6h_MF, x)[[1]] & Ri6h_vs_Rm6h_p.val<0.05 & Ri6h_vs_Rm6h_ratio < -2, Ri6h_vs_Rm6h_ratio])}, USE.NAMES = F))]

plot6h_full_MF$Term2 <- sapply(plot6h_full_MF$Term, CapStr, USE.NAMES = F)


pal<-brewer.pal(3,"PuOr")
data <- plot6h_full_MF[raw.p.value<0.05 & Positive+Negative>3]

limit <- round_any(max(data[,Positive])+max(data[,Negative]),5,ceiling)
sep <- if (limit>30) {10} else {5}

data <- data.table::melt(data = data, id.vars = c("Term2"), measure.vars = c("Positive", "Negative"), variable.name = "Fold", value.name = "Count")
data[Fold == "Negative", `:=` (Col = pal[3])]
data[Fold == "Positive", `:=` (Col = pal[1])]

plotMF6h <- ggplot() + geom_bar(data=data, aes(x = Term2, y=Count, fill=Col), position="stack", stat="identity") +
  geom_hline(yintercept = 0, color =c("white")) +
  scale_fill_identity("Number of modulated proteins", labels = c("Up accumulated", "Down accumulated"), breaks=c(pal[1],pal[3]), guide="legend") + 
  theme_hc() + 
  coord_flip() +
  labs(title="6hpi GO Molecular Function", y="",x="") +
  theme(plot.title = element_text(size=16, hjust=0.5)) +
  theme(axis.text = element_text(size=11)) +
  theme(axis.text.y = element_text(hjust=1)) +
  theme(legend.position = "bottom") +
  scale_y_continuous(breaks=seq(0,limit,sep), limits=c(0,limit)) +
  scale_x_discrete(label=function(x) stringr::str_trunc(x, 40))
plotMF6h
ggsave("MF6h.pdf", plot = plotMF6h, width = 10, height = 7, dpi = 600)
```


```{r}
diffGO6h_CC <- data.table(reducedTerms6h_CC)[go == parent, go]

plot6h_full_CC <- data.table(tab6h_CC)[GO.ID %in% diffGO6h_CC]

plot6h_full_CC[,`:=`(Positive = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata6h_CC, x)[[1]] & Ri6h_vs_Rm6h_p.val<0.05 & Ri6h_vs_Rm6h_ratio > 2, Ri6h_vs_Rm6h_ratio])}, USE.NAMES = F), Negative = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata6h_CC, x)[[1]] & Ri6h_vs_Rm6h_p.val<0.05 & Ri6h_vs_Rm6h_ratio < -2, Ri6h_vs_Rm6h_ratio])}, USE.NAMES = F))]

plot6h_full_CC$Term2 <- sapply(plot6h_full_CC$Term, CapStr, USE.NAMES = F)


pal<-brewer.pal(3,"PuOr")
data <- plot6h_full_CC[raw.p.value<0.05 & Positive+Negative>3]

limit <- round_any(max(data[,Positive])+max(data[,Negative]),5,ceiling)
sep <- if (limit>30) {10} else {5}

data <- data.table::melt(data = data, id.vars = c("Term2"), measure.vars = c("Positive", "Negative"), variable.name = "Fold", value.name = "Count")
data[Fold == "Negative", `:=` (Col = pal[3])]
data[Fold == "Positive", `:=` (Col = pal[1])]

plotCC6h <- ggplot() + geom_bar(data=data, aes(x = Term2, y=Count, fill=Col), position="stack", stat="identity") +
  geom_hline(yintercept = 0, color =c("white")) +
  scale_fill_identity("Number of modulated proteins", labels = c("Up accumulated", "Down accumulated"), breaks=c(pal[1],pal[3]), guide="legend") + 
  theme_hc() + 
  coord_flip() +
  labs(title="6hpi GO Celular Component", y="",x="") +
  theme(plot.title = element_text(size=16, hjust=0.5)) +
  theme(axis.text = element_text(size=11)) +
  theme(axis.text.y = element_text(hjust=1)) +
  theme(legend.position = "bottom") +
  scale_y_continuous(breaks=seq(0,limit,sep), limits=c(0,limit)) +
  scale_x_discrete(label=function(x) stringr::str_trunc(x, 40))
plotCC6h
ggsave("CC6h.pdf", plot = plotCC6h, width = 10, height = 7, dpi = 600)
```


```{r}
diffGO12h_BP <- data.table(reducedTerms12h_BP)[go == parent, go]

plot12h_full_BP <- data.table(tab12h_BP)[GO.ID %in% diffGO12h_BP]

plot12h_full_BP[,`:=`(Positive = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata12h_BP, x)[[1]] & Ri12h_vs_Rm12h_p.val<0.05 & Ri12h_vs_Rm12h_ratio > 2, Ri12h_vs_Rm12h_ratio])}, USE.NAMES = F), Negative = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata12h_BP, x)[[1]] & Ri12h_vs_Rm12h_p.val<0.05 & Ri12h_vs_Rm12h_ratio < -2, Ri12h_vs_Rm12h_ratio])}, USE.NAMES = F))]

plot12h_full_BP$Term2 <- sapply(plot12h_full_BP$Term, CapStr, USE.NAMES = F)


pal<-brewer.pal(3,"PuOr")
data <- plot12h_full_BP[raw.p.value<0.05 & Positive+Negative>3]

limit <- round_any(max(data[,Positive])+max(data[,Negative]),5,ceiling)
sep <- if (limit>30) {10} else {5}

data <- data.table::melt(data = data, id.vars = c("Term2"), measure.vars = c("Positive", "Negative"), variable.name = "Fold", value.name = "Count")
data[Fold == "Negative", `:=` (Col = pal[3])]
data[Fold == "Positive", `:=` (Col = pal[1])]

plotBP12h <- ggplot() + geom_bar(data=data, aes(x = Term2, y=Count, fill=Col), position="stack", stat="identity") +
  geom_hline(yintercept = 0, color =c("white")) +
  scale_fill_identity("Number of modulated proteins", labels = c("Up accumulated", "Down accumulated"), breaks=c(pal[1],pal[3]), guide="legend") + 
  theme_hc() + 
  coord_flip() +
  labs(title="12hpi GO Biological Process", y="",x="") +
  theme(plot.title = element_text(size=16, hjust=0.5)) +
  theme(axis.text = element_text(size=11)) +
  theme(axis.text.y = element_text(hjust=1)) +
  theme(legend.position = "bottom") +
  scale_y_continuous(breaks=seq(0,limit,sep), limits=c(0,limit)) +
  scale_x_discrete(label=function(x) stringr::str_trunc(x, 40))
plotBP12h
ggsave("BP12h.pdf", plot = plotBP12h, width = 10, height = 7, dpi = 600)
```


```{r}
diffGO12h_MF <- data.table(reducedTerms12h_MF)[go == parent, go]

plot12h_full_MF <- data.table(tab12h_MF)[GO.ID %in% diffGO12h_MF]

plot12h_full_MF[,`:=`(Positive = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata12h_MF, x)[[1]] & Ri12h_vs_Rm12h_p.val<0.05 & Ri12h_vs_Rm12h_ratio > 2, Ri12h_vs_Rm12h_ratio])}, USE.NAMES = F), Negative = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata12h_MF, x)[[1]] & Ri12h_vs_Rm12h_p.val<0.05 & Ri12h_vs_Rm12h_ratio < -2, Ri12h_vs_Rm12h_ratio])}, USE.NAMES = F))]

plot12h_full_MF$Term2 <- sapply(plot12h_full_MF$Term, CapStr, USE.NAMES = F)


pal<-brewer.pal(3,"PuOr")
data <- plot12h_full_MF[raw.p.value<0.05 & Positive+Negative>3]

limit <- round_any(max(data[,Positive])+max(data[,Negative]),5,ceiling)
sep <- if (limit>30) {10} else {5}

data <- data.table::melt(data = data, id.vars = c("Term2"), measure.vars = c("Positive", "Negative"), variable.name = "Fold", value.name = "Count")
data[Fold == "Negative", `:=` (Col = pal[3])]
data[Fold == "Positive", `:=` (Col = pal[1])]

plotMF12h <- ggplot() + geom_bar(data=data, aes(x = Term2, y=Count, fill=Col), position="stack", stat="identity") +
  geom_hline(yintercept = 0, color =c("white")) +
  scale_fill_identity("Number of modulated proteins", labels = c("Up accumulated", "Down accumulated"), breaks=c(pal[1],pal[3]), guide="legend") + 
  theme_hc() + 
  coord_flip() +
  labs(title="12hpi GO Molecular Function", y="",x="") +
  theme(plot.title = element_text(size=16, hjust=0.5)) +
  theme(axis.text = element_text(size=11)) +
  theme(axis.text.y = element_text(hjust=1)) +
  theme(legend.position = "bottom") +
  scale_y_continuous(breaks=seq(0,limit,sep), limits=c(0,limit)) +
  scale_x_discrete(label=function(x) stringr::str_trunc(x, 40))
plotMF12h
ggsave("MF12h.pdf", plot = plotMF12h, width = 10, height = 7, dpi = 600)
```


```{r}
diffGO12h_CC <- data.table(reducedTerms12h_CC)[go == parent, go]

plot12h_full_CC <- data.table(tab12h_CC)[GO.ID %in% diffGO12h_CC]

plot12h_full_CC[,`:=`(Positive = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata12h_CC, x)[[1]] & Ri12h_vs_Rm12h_p.val<0.05 & Ri12h_vs_Rm12h_ratio > 2, Ri12h_vs_Rm12h_ratio])}, USE.NAMES = F), Negative = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata12h_CC, x)[[1]] & Ri12h_vs_Rm12h_p.val<0.05 & Ri12h_vs_Rm12h_ratio < -2, Ri12h_vs_Rm12h_ratio])}, USE.NAMES = F))]

plot12h_full_CC$Term2 <- sapply(plot12h_full_CC$Term, CapStr, USE.NAMES = F)


pal<-brewer.pal(3,"PuOr")
data <- plot12h_full_CC[raw.p.value<0.05 & Positive+Negative>3]

limit <- round_any(max(data[,Positive])+max(data[,Negative]),5,ceiling)
sep <- if (limit>30) {10} else {5}

data <- data.table::melt(data = data, id.vars = c("Term2"), measure.vars = c("Positive", "Negative"), variable.name = "Fold", value.name = "Count")
data[Fold == "Negative", `:=` (Col = pal[3])]
data[Fold == "Positive", `:=` (Col = pal[1])]

plotCC12h <- ggplot() + geom_bar(data=data, aes(x = Term2, y=Count, fill=Col), position="stack", stat="identity") +
  geom_hline(yintercept = 0, color =c("white")) +
  scale_fill_identity("Number of modulated proteins", labels = c("Up accumulated", "Down accumulated"), breaks=c(pal[1],pal[3]), guide="legend") + 
  theme_hc() + 
  coord_flip() +
  labs(title="12hpi GO Celular Component", y="",x="") +
  theme(plot.title = element_text(size=16, hjust=0.5)) +
  theme(axis.text = element_text(size=11)) +
  theme(axis.text.y = element_text(hjust=1)) +
  theme(legend.position = "bottom") +
  scale_y_continuous(breaks=seq(0,limit,sep), limits=c(0,limit)) +
  scale_x_discrete(label=function(x) stringr::str_trunc(x, 40))
plotCC12h
ggsave("CC12h.pdf", plot = plotCC12h, width = 10, height = 7, dpi = 600)
```

```{r}
final_plot <- ggarrange(plotBP6h + theme(legend.title = element_text(size = 16), legend.text = element_text(size = 16)), 
                        plotMF6h, 
                        plotCC6h, 
                        plotBP12h, 
                        plotMF12h, 
                        plotCC12h,
                        nrow = 2, ncol = 3, common.legend = TRUE, legend = "bottom")

final_plot
ggsave("GO_Enrichment.pdf", plot = final_plot, width = 20, height = 14, dpi = 300)
```

# Save Go Results

Write GO results to Excel sheets

```{r}
writeGO2Excel <- function(fileName, goresults, topGOdata, dep_results, FC){
  upreg <- paste0(">", log2(FC))
  downreg <- paste0("<-", log2(FC))
  negStyle <- createStyle(fontColour = "#9C0006", bgFill = "#FFC7CE")
  posStyle <- createStyle(fontColour = "#006100", bgFill = "#C6EFCE")
  
  excel <- createWorkbook()

  addWorksheet(excel, "Summary")
  writeData(goresults[goresults$raw.p.value < 0.05, ], wb = excel, sheet = "Summary", rowNames = FALSE)

  GoDF <- list()
  for (x in 1:sum(goresults$raw.p.value < 0.05)) {
    goID <- goresults[x, "GO.ID"]
    ids <- genesInTerm(topGOdata, goID)
    annot <- data.table(AnnotationDbi::select(org.Vvinifera.eg.db, keys=ids[[1]], column=c("symbol", "gene_name"), keytype="gene_id"))
    tmp <- data.table(dep_results)
    expVals <- tmp[GENE_GI %in% annot$gene_id, ]
    final_annot <- merge.data.table(annot, expVals, by.x ="gene_id", by.y = "GENE_GI")
    setorder(final_annot, Ri6h_vs_Rm6h_p.val)
    GoDF[[goID]] <- final_annot
  }

  for (x in 1:sum(goresults$raw.p.value < 0.05)) {
    goID <- goresults[x, "GO.ID"]
    sheetname <- substring(goresults[x, "GO.ID"], 4)
    addWorksheet(excel, sheetname)
    writeData(GoDF[[x]], wb = excel, sheet = sheetname, rowNames = FALSE)
    writeFormula(excel, "Summary", startRow = x+1, 
                 x = makeHyperlinkString(sheet = sheetname, row = 1, col = 1
                                         , text = goID))
    conditionalFormatting(excel, sheetname, cols=14, rows=2:(nrow(GoDF[[x]])+1), rule=upreg, style = posStyle)
    conditionalFormatting(excel, sheetname, cols=14, rows=2:(nrow(GoDF[[x]])+1), rule=downreg, style = posStyle)
    conditionalFormatting(excel, sheetname, cols=13, rows=2:(nrow(GoDF[[x]])+1), rule=upreg, style = posStyle)
    conditionalFormatting(excel, sheetname, cols=13, rows=2:(nrow(GoDF[[x]])+1), rule=downreg, style = posStyle)
    conditionalFormatting(excel, sheetname, cols=6, rows=2:(nrow(GoDF[[x]])+1), rule="<0.05", style = posStyle)
    conditionalFormatting(excel, sheetname, cols=7, rows=2:(nrow(GoDF[[x]])+1), rule="<0.05", style = posStyle)
  }
  
  saveWorkbook(excel, file = fileName, overwrite = TRUE)
}
```

```{r, message=FALSE}
writeGO2Excel("results/6hpi_GO_BP.xlsx", tab6h_BP, GOdata6h_BP, data_results, 2)
writeGO2Excel("results/6hpi_GO_MF.xlsx", tab6h_MF, GOdata6h_MF, data_results, 2)
writeGO2Excel("results/6hpi_GO_CC.xlsx", tab6h_CC, GOdata6h_CC, data_results, 2)
writeGO2Excel("results/12hpi_GO_BP.xlsx", tab12h_BP, GOdata12h_BP, data_results, 2)
writeGO2Excel("results/12hpi_GO_MF.xlsx", tab12h_MF, GOdata12h_MF, data_results, 2)
writeGO2Excel("results/12hpi_GO_CC.xlsx", tab12h_CC, GOdata12h_CC, data_results, 2)
```

Save timepoints

```{r}
signif.6hpi <- data_results2[abs(Ri6h_vs_Rm6h_ratio) > 2 & Ri6h_vs_Rm6h_p.val < 0.05,
              c("PROTEIN_NAME","ID","PROTEIN_ID","Ri6h_vs_Rm6h_p.val","Ri6h_vs_Rm6h_ratio","GENE_GI","name")]
signif.12hpi <- data_results2[abs(Ri12h_vs_Rm12h_ratio) > 2 & Ri12h_vs_Rm12h_p.val < 0.05,
              c("PROTEIN_NAME","ID","PROTEIN_ID","Ri12h_vs_Rm12h_p.val","Ri12h_vs_Rm12h_ratio","GENE_GI","name")]

write.csv(signif.6hpi, "significant_6hpi.csv", row.names = F)
write.csv(signif.12hpi, "significant_12hpi.csv", row.names = F)
```

```{r}
down.AT <- data_results2[(Ri12h_vs_Rm12h_ratio - Ri6h_vs_Rm6h_ratio) < -2 & Ri6h_vs_Rm6h_p.val < 0.05 & Ri12h_vs_Rm12h_p.val < 0.05,
              c("PROTEIN_NAME","ID","PROTEIN_ID","Ri6h_vs_Rm6h_p.val","Ri12h_vs_Rm12h_p.val","Ri6h_vs_Rm6h_ratio","Ri12h_vs_Rm12h_ratio","GENE_GI","name")]
up.AT <- data_results2[(Ri12h_vs_Rm12h_ratio - Ri6h_vs_Rm6h_ratio) > 2 & Ri6h_vs_Rm6h_p.val < 0.05 & Ri12h_vs_Rm12h_p.val < 0.05,
              c("PROTEIN_NAME","ID","PROTEIN_ID","Ri6h_vs_Rm6h_p.val","Ri12h_vs_Rm12h_p.val","Ri6h_vs_Rm6h_ratio","Ri12h_vs_Rm12h_ratio","GENE_GI","name")]
write.csv(down.AT, "down_along_time.csv", row.names = F)
write.csv(up.AT, "up_along_time.csv", row.names = F)
```

Write tables for manual curation

```{r}
diffGO6h_BP <- data.table(reducedTerms6h_BP)[, go]

plot6h_full_BP <- data.table(tab6h_BP)[GO.ID %in% diffGO6h_BP]

plot6h_full_BP[,`:=`(Positive = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata6h_BP, x)[[1]] & Ri6h_vs_Rm6h_p.val<0.05 & Ri6h_vs_Rm6h_ratio > 2, Ri6h_vs_Rm6h_ratio])}, USE.NAMES = F), Negative = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata6h_BP, x)[[1]] & Ri6h_vs_Rm6h_p.val<0.05 & Ri6h_vs_Rm6h_ratio < -2, Ri6h_vs_Rm6h_ratio])}, USE.NAMES = F))]

plot6h_full_BP$Term2 <- sapply(plot6h_full_BP$Term, CapStr, USE.NAMES = F)

BP_6hpi <- merge.data.table(plot6h_full_BP,reducedTerms6h_BP, by.x = "GO.ID", by.y = "go")
write.csv(BP_6hpi, "BP_6hpi.csv", row.names = FALSE)


diffGO6h_MF <- data.table(reducedTerms6h_MF)[, go]

plot6h_full_MF <- data.table(tab6h_MF)[GO.ID %in% diffGO6h_MF]

plot6h_full_MF[,`:=`(Positive = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata6h_MF, x)[[1]] & Ri6h_vs_Rm6h_p.val<0.05 & Ri6h_vs_Rm6h_ratio > 2, Ri6h_vs_Rm6h_ratio])}, USE.NAMES = F), Negative = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata6h_MF, x)[[1]] & Ri6h_vs_Rm6h_p.val<0.05 & Ri6h_vs_Rm6h_ratio < -2, Ri6h_vs_Rm6h_ratio])}, USE.NAMES = F))]

plot6h_full_MF$Term2 <- sapply(plot6h_full_MF$Term, CapStr, USE.NAMES = F)

MF_6hpi <- merge.data.table(plot6h_full_MF,reducedTerms6h_MF, by.x = "GO.ID", by.y = "go")
write.csv(MF_6hpi, "MF_6hpi.csv", row.names = FALSE)


diffGO6h_CC <- data.table(reducedTerms6h_CC)[, go]

plot6h_full_CC <- data.table(tab6h_CC)[GO.ID %in% diffGO6h_CC]

plot6h_full_CC[,`:=`(Positive = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata6h_CC, x)[[1]] & Ri6h_vs_Rm6h_p.val<0.05 & Ri6h_vs_Rm6h_ratio > 2, Ri6h_vs_Rm6h_ratio])}, USE.NAMES = F), Negative = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata6h_CC, x)[[1]] & Ri6h_vs_Rm6h_p.val<0.05 & Ri6h_vs_Rm6h_ratio < -2, Ri6h_vs_Rm6h_ratio])}, USE.NAMES = F))]

plot6h_full_CC$Term2 <- sapply(plot6h_full_CC$Term, CapStr, USE.NAMES = F)

CC_6hpi <- merge.data.table(plot6h_full_CC,reducedTerms6h_CC, by.x = "GO.ID", by.y = "go")
write.csv(CC_6hpi, "CC_6hpi.csv", row.names = FALSE)



diffGO12h_BP <- data.table(reducedTerms12h_BP)[, go]

plot12h_full_BP <- data.table(tab12h_BP)[GO.ID %in% diffGO12h_BP]

plot12h_full_BP[,`:=`(Positive = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata12h_BP, x)[[1]] & Ri12h_vs_Rm12h_p.val<0.05 & Ri12h_vs_Rm12h_ratio > 2, Ri12h_vs_Rm12h_ratio])}, USE.NAMES = F), Negative = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata12h_BP, x)[[1]] & Ri12h_vs_Rm12h_p.val<0.05 & Ri12h_vs_Rm12h_ratio < -2, Ri12h_vs_Rm12h_ratio])}, USE.NAMES = F))]

plot12h_full_BP$Term2 <- sapply(plot12h_full_BP$Term, CapStr, USE.NAMES = F)

BP_12hpi <- merge.data.table(plot12h_full_BP,reducedTerms12h_BP, by.x = "GO.ID", by.y = "go")
write.csv(BP_12hpi, "BP_12hpi.csv", row.names = FALSE)


diffGO12h_MF <- data.table(reducedTerms12h_MF)[, go]

plot12h_full_MF <- data.table(tab12h_MF)[GO.ID %in% diffGO12h_MF]

plot12h_full_MF[,`:=`(Positive = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata12h_MF, x)[[1]] & Ri12h_vs_Rm12h_p.val<0.05 & Ri12h_vs_Rm12h_ratio > 2, Ri12h_vs_Rm12h_ratio])}, USE.NAMES = F), Negative = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata12h_MF, x)[[1]] & Ri12h_vs_Rm12h_p.val<0.05 & Ri12h_vs_Rm12h_ratio < -2, Ri12h_vs_Rm12h_ratio])}, USE.NAMES = F))]

plot12h_full_MF$Term2 <- sapply(plot12h_full_MF$Term, CapStr, USE.NAMES = F)

MF_12hpi <- merge.data.table(plot12h_full_MF,reducedTerms12h_MF, by.x = "GO.ID", by.y = "go")
write.csv(MF_12hpi, "MF_12hpi.csv", row.names = FALSE)


diffGO12h_CC <- data.table(reducedTerms12h_CC)[, go]

plot12h_full_CC <- data.table(tab12h_CC)[GO.ID %in% diffGO12h_CC]

plot12h_full_CC[,`:=`(Positive = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata12h_CC, x)[[1]] & Ri12h_vs_Rm12h_p.val<0.05 & Ri12h_vs_Rm12h_ratio > 2, Ri12h_vs_Rm12h_ratio])}, USE.NAMES = F), Negative = sapply(GO.ID, function(x){length(final_results[GENE_GI %in% genesInTerm(object = GOdata12h_CC, x)[[1]] & Ri12h_vs_Rm12h_p.val<0.05 & Ri12h_vs_Rm12h_ratio < -2, Ri12h_vs_Rm12h_ratio])}, USE.NAMES = F))]

plot12h_full_CC$Term2 <- sapply(plot12h_full_CC$Term, CapStr, USE.NAMES = F)

CC_12hpi <- merge.data.table(plot12h_full_CC,reducedTerms12h_CC, by.x = "GO.ID", by.y = "go")
write.csv(CC_12hpi, "CC_12hpi.csv", row.names = FALSE)
```

