---
title: "Prometeus MaxQuanta Data Analisys"
author: "Rui Nascimento"
date: "`r format(Sys.time(), '%d %B, %Y (%X)')`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Vitis Proteome Analysis

Repository of code for Data Analysis of MaxQuant results of LC-MS/MS material.

Using [Proteus](https://github.com/bartongroup/Proteus) R package.

### The data:

* _Vitis vinifera_ leaf Proteome.
* `Control` and `Inoculated` with _P. viticola_.
* At 3 time points, 6, 12 and 24 hpi.

# Prometeus Pipeline

### Disclamer

I will be trying to use the Proteus pipeline using just the `protein groups file` (without the Evidence file), it will be missing peptide information for sure and maybe some other relevant information.

## Load required libraries

```{r libraryLoad, warning=FALSE}
library(limma)
library(proteus)
```

## Metadata object

Create a metadata object consistent with our dataset

```{r}
metadataFile <- "metadata.csv"
meta <- read.delim(metadataFile, header=TRUE, sep=",")
meta
```

## Read the protein groups file

```{r}
proteinGroupsFile <- "Vitis_Protein_Groups.tsv"
prot.MQ <- readProteinGroups(proteinGroupsFile, meta)
```

## Data Normalization

Normalize datasets.

```{r}
prot.MQ.med <- normalizeData(prot.MQ)
prot.MQ.quant <- normalizeData(prot.MQ, norm.fun=limma::normalizeQuantiles)
```

```{r}
plotSampleDistributions(prot.MQ, title="Not normalized", fill="condition", method="violin")

plotSampleDistributions(prot.MQ.med, title="Median normalization", fill="condition", method="violin")

plotSampleDistributions(prot.MQ.quant, title="Quantile normalization", fill="condition", method="violin")
```

The best method looks like the quantile normalization and will be used for the rest of the analysis.

## Differential expression at 6hpi

```{r}
res.6hpi <- limmaDE(prot.MQ.quant, sig.level=0.05, conditions = c("Ri6h", "Rm6h"))
```

### List of significantly differentially accomulated proteins at 6hpi

(Only showing the first 5 for space constrains)
```{r}
head(res.6hpi[which(res.6hpi$significant), c("protein", "logFC", "adj.P.Val")])
```


### Proteins in only one condition

* Proteins only in Inoculated leafs
```{r}
only.Ri6h <- which(prot.MQ.quant$detect$`Ri6h` & !prot.MQ.quant$detect$Rm6h)
#as.character(prot.MQ.quant$proteins[only.Ri6h])
```
There were detected `r length(only.Ri6h)` proteins in at least 1 replicate in Inoculated samples but not detected in Mock samples.


* Proteins only in Mock leafs
```{r}
only.Rm6h <- which(!prot.MQ.quant$detect$`Ri6h` & prot.MQ.quant$detect$Rm6h)
# as.character(prot.MQ.quant$proteins[only.Rm6h])
```

There were detected `r length(only.Rm6h)` proteins in at least 1 replicate in Mock samples but not detected in Inoculated samples.

### Data Visualization

```{r}
plotFID(prot.MQ.quant, pair = c("Ri6h", "Rm6h"))
```

```{r}
plotVolcano(res.6hpi)
```

```{r}
plotPdist(res.6hpi)
```


## Differential expression at 12hpi

```{r}
res.12hpi <- limmaDE(prot.MQ.quant, sig.level=0.05, conditions = c("Ri12h", "Rm12h"))
```

### List of significantly differentially accomulated proteins at 12hpi

(Only showing the first 5 for space constrains)
```{r}
head(res.12hpi[which(res.12hpi$significant), c("protein", "logFC", "adj.P.Val")])
```


### Proteins in only one condition

* Proteins only in Inoculated leafs
```{r}
only.Ri12h <- which(prot.MQ.quant$detect$`Ri12h` & !prot.MQ.quant$detect$Rm12h)
#as.character(prot.MQ.quant$proteins[only.Ri12h])
```
There were detected `r length(only.Ri12h)` proteins in at least 1 replicate in Inoculated samples but not detected in Mock samples.


* Proteins only in Mock leafs
```{r}
only.Rm12h <- which(!prot.MQ.quant$detect$`Ri12h` & prot.MQ.quant$detect$Rm12h)
# as.character(prot.MQ.quant$proteins[only.Rm12h])
```

There were detected `r length(only.Rm12h)` proteins in at least 1 replicate in Mock samples but not detected in Inoculated samples.

### Data Visualization

```{r}
plotFID(prot.MQ.quant, pair = c("Ri12h", "Rm12h"))
```

```{r}
plotVolcano(res.12hpi)
```

```{r}
plotPdist(res.12hpi)
```


## Differential expression at 24hpi

```{r}
res.24hpi <- limmaDE(prot.MQ.quant, sig.level=0.05, conditions = c("Ri24h", "Rm24h"))
```

### List of significantly differentially accomulated proteins at 24hpi

(Only showing the first 5 for space constrains)
```{r}
head(res.24hpi[which(res.24hpi$significant), c("protein", "logFC", "adj.P.Val")])
```


### Proteins in only one condition

* Proteins only in Inoculated leafs
```{r}
only.Ri24h <- which(prot.MQ.quant$detect$`Ri24h` & !prot.MQ.quant$detect$Rm24h)
#as.character(prot.MQ.quant$proteins[only.Ri24h])
```
There were detected `r length(only.Ri24h)` proteins in at least 1 replicate in Inoculated samples but not detected in Mock samples.


* Proteins only in Mock leafs
```{r}
only.Rm24h <- which(!prot.MQ.quant$detect$`Ri24h` & prot.MQ.quant$detect$Rm24h)
# as.character(prot.MQ.quant$proteins[only.Rm24h])
```

There were detected `r length(only.Rm24h)` proteins in at least 1 replicate in Mock samples but not detected in Inoculated samples.

### Data Visualization

```{r}
plotFID(prot.MQ.quant, pair = c("Ri24h", "Rm24h"))
```

```{r}
plotVolcano(res.24hpi)
```

```{r}
plotPdist(res.24hpi)
```

I can also plot the results for individual proteins and peptides with the Evidence file...

And also Interactive Plots locally.


Save results to file:
```{r}
write.table(res.6hpi, "6hpi_DE_output.tsv", sep = "\t", row.names = FALSE)
write.table(res.12hpi, "12hpi_DE_output.tsv", sep = "\t", row.names = FALSE)
write.table(res.24hpi, "24hpi_DE_output.tsv", sep = "\t", row.names = FALSE)
```

