---
title: "Proteus MaxQuant Data Analisys"
author: "Rui Nascimento"
date: "`r format(Sys.time(), '%d %B, %Y (%X)')`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Vitis Proteome Analysis

Repository of code for Data Analysis of MaxQuant results of LC-MS/MS material.

Using [Proteus](https://github.com/bartongroup/Proteus) R package.

### The data:

* _Vitis vinifera_ leaf Proteome.
* `Control` and `Inoculated` with _P. viticola_.
* At 3 time points, 6, 12 and 24 hpi.

# Proteus Pipeline

## Load required libraries

```{r libraryLoad, warning=FALSE}
library(limma)
library(proteus)
```

## Load Metadata object

Create a metadata object consistent with our dataset

```{r}
metadataFile <- "metadata.csv"
meta <- read.delim(metadataFile, header=TRUE, sep=",")
meta
```

## Read evidence file

```{r}
evidenceFile <- "evidence.txt"
evi <- readEvidenceFile(evidenceFile)
head(evi)
```

## Peptide Data

```{r}
pepdat <- makePeptideTable(evi, meta)
pepdat$tab[1:5, 1:18]
summary(pepdat)
```


## Number of peptides

```{r}
plotCount(pepdat)
```

## Jaccard similarity

```{r}
plotDetectionSimilarity(pepdat, bin.size = 0.02)
```

## Clustering

```{r}
plotClustering(pepdat)
```

## Protein Data

Create the protein dataset

```{r}
prodat <- makeProteinTable(pepdat)
summary(prodat)
```

## Normalization

Non normalized samples

```{r}
plotSampleDistributions(prodat, title="Not normalized", fill="condition", method="violin")
```

Median Normalized Samples
```{r}
prodat.med <- normalizeData(prodat)
plotSampleDistributions(prodat.med, title="Median normalization", fill="condition", method="violin")
```

Quantile Normalized Samples
```{r}
prodat.quant <- normalizeData(prodat, norm.fun=limma::normalizeQuantiles)
plotSampleDistributions(prodat.quant, title="Quantile normalization", fill="condition", method="violin")
```

## Clustering

```{r}
plotClustering(prodat.quant)
```

## MDS Plot of normalized data

```{r}
plotMDS(prodat.quant$tab, col = as.numeric(factor(meta$condition)), labels = meta$condition)
```

## Read Maxquant Protein groups

```{r}
proteinGroupsFile <- "Vitis_Protein_Groups.tsv"
prot.MQ <- readProteinGroups(proteinGroupsFile, meta)
```


## Limma DE analysis

```{r}
res6 <- limmaDE(prodat.med, sig.level=0.05, conditions = c("Ri6h", "Rm6h"))
subset(res6, significant == "TRUE")
```

```{r}
res12 <- limmaDE(prodat.med, sig.level=0.05, conditions = c("Ri12h", "Rm12h"))
subset(res12, significant == "TRUE")
```


```{r}
res24 <- limmaDE(prodat.med, sig.level=0.05, conditions = c("Ri24h", "Rm24h"))
subset(res24, significant == "TRUE")
```
